---
title: ''
output:
  html_document:
    df_print: paged
  pdf_document: default
---
#  Diff between Rebecca Turner's taxonomy file and the authoritative taxonomy table
#  August 14, 2020


```{r, message = FALSE}
library(tidyverse) ; library(testthat)
```

```{r}
# Rebecca's file
rt_aug_file <- read_csv("nfs_data/data/clean_data/dated_tax_files/taxonomy_table2020Aug12b.csv")
rt_aug_file <- rt_aug_file %>% arrange(taxon_id) %>% 
               mutate_all(~gsub("(*UCP)\\s\\+|\\W+$", "", . , perl=TRUE)) %>%
               mutate(taxon_id = as.double(taxon_id), 
                      usagekey = as.numeric(usagekey),
                      synonym = as.logical(synonym), 
                      acceptedusagekey = as.numeric(acceptedusagekey))

# taxonomy table
tax_table <- read_csv("nfs_data/data/clean_data/taxonomy_table.csv")
tax_table <- tax_table %>% arrange(taxon_id)

nrow(rt_aug_file) ; nrow(tax_table)  #  rows difference
names(rt_aug_file) ; names(tax_table) # column names difference
```


# Diff between the two files:

```{r}
anti_join(tax_table, rt_aug_file, by = "taxon_id")  # only diff by taxon_id

diff3 <- anti_join(tax_table, rt_aug_file)  # returns differences
str(diff3)
```


# I suspect that taxa are being numbered differently between the two cleaning codes, so testing that here:
```{r}
# pasting the taxon_id and user_supplied_name columns together to 
tax_table2 <- tax_table %>% mutate(id_usr_supp_nm = paste(taxon_id, user_supplied_name))

rt_aug_file2 <- rt_aug_file %>% mutate(id_usr_supp_nm = paste(taxon_id, user_supplied_name))
```

# Yes, there are many rows with different taxon IDs!!
```{r}
 # rows of tax_table2 that don't have a match in rt_aug_file2 based on the pasted-together column
diff4 <- anti_join(tax_table2, rt_aug_file2, by = "id_usr_supp_nm")
str(diff4)

# rows of rt_aug_file2 that don't have a match in tax_table2
# diff5 <- anti_join(rt_aug_file2, tax_table2, by = "id_usr_supp_nm") 
# str(diff5)
```
### Not sure how to fix the issue of differing taxon IDs.  Will talk to Rebecca about this and try to resolve in the taxonomic cleaning code.


# But there are still other rows that still don't match!!
```{r}
# Subset diff3 to remove rows with taxon_id mismatches, leaving all other mismatches
other_probs <- anti_join(diff3, diff4) %>% arrange(user_supplied_name) %>% 
               arrange(user_supplied_name)

# get a vector of these taxa
usr_nms_diff <- other_probs$user_supplied_name

# subset Rebecca's file with these still-mismatching rows
rt_diff <- rt_aug_file %>% filter(user_supplied_name %in% usr_nms_diff) %>% 
           arrange(user_supplied_name)
```

# This is actually quite useful!  Gives a print out of which columns have issues and how many issues.
```{r, error = TRUE}
testthat::expect_equal(other_probs, rt_diff)   
```

# further investigation by column based on testthat results
```{r}
column_na_test <- function(column_nm){
                  tt_df <- other_probs[is.na(other_probs[[column_nm]]), ]
                  rt_df <- rt_diff[is.na(rt_diff[[column_nm]]), ]
               
                  aj_diff <- anti_join(tt_df, rt_df, by = "user_supplied_name")
                  
                  aj_diff2 <- anti_join(rt_df, tt_df, by = "user_supplied_name")
                    
                  ifelse(nrow(aj_diff) == 0, return_df <- aj_diff2, return_df <- aj_diff)
                 
                  return(return_df)
                  }

```

```{r}
column_na_test("status")
```

```{r}
column_na_test("matchtype")
```

```{r}
column_na_test("usagekey")
```

```{r}
column_na_test("family")
```

```{r}
column_na_test("genus")
```

```{r}
sp_test <- column_na_test("species")
nrow(sp_test)
sp_test
```

```{r}
anti_join(other_probs, rt_diff, by = "synonym")
```

```{r}
aj_gs <- anti_join(other_probs, rt_diff, by = "genus_species")
aj_gs
```

### Note: I'm disregarding the 328 mismatches in taxonomic_authority column, as those are not critical


# Check Occurrence table for taxon ID matches
```{r}
# Read in the occurrance table:
occ_tab <- read.csv("nfs_data/data/clean_data/occurrence_table.csv", stringsAsFactors=F)  # read in the occurrence table

```

```{r}
# random subset to check matching
tt <- tax_table %>% filter(taxon_id %in% 500:520) # subset taxonomy table

oc <- occ_tab %>% filter(taxon_id %in% 500:520) # subset occurrence table

nrow(tt) ; nrow(oc)

```

```{r}
anti_join(tt, oc)

```

```{r, error = TRUE}
expect_equal(tt,oc)
```

```{r}
tt
```

```{r}
oc
```

### Taxon IDs match between taxonomy table and occurrence table


# working with Rebecca Aug 25, 2020

```{r}
# occurrence table
occ_table <- read_csv("nfs_data/data/clean_data/occurrence_table.csv")
occ_table <- occ_table %>% arrange(user_supplied_name)

```

```{r}
tt_sub <- tax_table %>% select(user_supplied_name, rank, order, family, genus_species)

rt_aug_sub <- rt_aug_file %>% select(user_supplied_name, rank, order, family, genus_species)
```

## Looking at differences in user_supplied_name columns
```{r}
usn <- setdiff(tt_sub$user_supplied_name, rt_aug_sub$user_supplied_name)

usn_r <- setdiff(rt_aug_sub$user_supplied_name, tt_sub$user_supplied_name)

```
## There are encoding problems with spaces and parentheses!  


### Comparing families for Rebecca's stuff, especially Coleoptera
```{r}
tt_sub2 <- tax_table %>% select(rank, family, genus_species, user_supplied_name) %>% filter(rank == "species")
rt2 <- rt_aug_file %>% select(rank, family , genus_species, user_supplied_name) %>% filter(rank == "species")

fj2 <- full_join(tt_sub2, rt2, by = "genus_species") 

```

```{r}
fx <- fj2[!(fj2$family.x == fj2$family.y),]

```

#  Trying to figure out encoding differences - 
```{r}
# 1) Reading in with read.csv() vs. read_csv() doesn't change the character encoding issues.

# 2) 
# Quentin's work-around function  - it works well on spaces!! :-)   
read.csv.better <- function(file_path, ...) {
                   read.csv_args <- list(...)
                   txt <- system2("perl", args = paste("-CSDA -plE \"s/\\s/ /g\"",
                                                       file_path), stdout = TRUE)
                   conn <- textConnection(txt) 
                   do.call(read.csv, args = c(list(file = conn), read.csv_args))
                   }

rt_s1 <- read.csv.better("nfs_data/data/clean_data/dated_tax_files/taxonomy_table2020Aug12b.csv", stringsAsFactors = FALSE) 

tt_sub <- tax_table %>% select(user_supplied_name, rank, order, family, genus_species)

usn_b1 <- setdiff(rt_s1$user_supplied_name, tt_sub$user_supplied_name)
sort(usn_b1)

usn_b2 <- setdiff(tt_sub$user_supplied_name, rt_s1$user_supplied_name)
sort(usn_b2)
```

```{r}
# 3) determine character encoding
readr::guess_encoding("nfs_data/data/clean_data/dated_tax_files/taxonomy_table2020Aug12b.csv")  # useful list of encodings for file
readr::guess_encoding("nfs_data/data/clean_data/taxonomy_table.csv")

# read.csv(file_to_read, header=TRUE, sep=",", fileEncoding=file_encoding, dec=".")

rt_enc2 <- read.csv("nfs_data/data/clean_data/dated_tax_files/taxonomy_table2020Aug12b.csv", fileEncoding = "UTF-8", dec=".")  # doesn't fix parentheses encoding
rt_enc2$user_supplied_name <- enc2utf8(rt_enc2$user_supplied_name)  # doesn't solve the encoding problem with parentheses
          
tt_sub <- tax_table %>% select(user_supplied_name, rank, order, family, genus_species)

usn_r2 <- setdiff(rt_enc2$user_supplied_name, tt_sub$user_supplied_name)
sort(usn_r2)

usn2 <- setdiff(tt_sub$user_supplied_name, rt_enc2$user_supplied_name)
sort(usn2)
```







