---
title: "README for Insect Invasions Taxonomic Tables"
output: pdf_document
---

```{r libs, echo=FALSE, message=FALSE}
# NOTE: Do not Knit this file normally.  Just run the last code chunk to knit a PDF to a different location.  

knitr::opts_chunk$set(comment = NA)
library(ezknitr); library(rmarkdown); library(taxize); library(dplyr)

in_dir <- "~/insect_invasions/example_code"
out_dir <- "/nfs/insectinvasions-data/data/clean_data"
```

# Last Update to Taxonomic Files: 
```{r dates, echo=FALSE}
tab_loc <- file.path("/nfs/insectinvasions-data/data/clean_data")
tables <- dir(tab_loc, pattern = "^[a-z]+_table", full.names=TRUE)
file.info(tables, extra_cols = FALSE)[,c(1,5)]
```

## Reason for last update:
Updates to Australian and North American raw datasets, and some new corrections in the fixed file were incorporated.  Re-ran all `make_XXX_table.R` scripts to update tables.  

## Taxonomic Database Citations
```{r citations, echo=FALSE}
ctime1 <- as.character(file.info(tables, extra_cols = FALSE)[3,c(5)])
gbif_access_date <- stringr::str_extract(ctime1, "^.{10}")
access_year <- stringr::str_extract(ctime1, "^.{4}")
```

```{r gbif, echo=FALSE, results='asis'}
cat(paste0("GBIF Secretariat ", access_year,". GBIF Backbone Taxonomy. Checklist dataset https://doi.org/10.15468/39omei accessed via GBIF.org on ",
           gbif_access_date,"."))
```

```{r other_databases, echo=FALSE}
db_list <- c("Catalogue of Life", "Wikispecies", "ITIS", "NCBI", "The Interim Register of Marine and Nonmarine Genera", "	Encyclopedia of Life",
             "Aphid Species File", "Index to Organism Names", "uBio NameBank")
db_info <- subset(gnr_datasources(), title %in% db_list) 
print.data.frame(db_info %>% select(id, title, updated_at), row.names = FALSE) 
```


## Raw Data Files
Raw data files are stored on the shared data directory at SESYNC (`/nfs/insectinvasions-data` or [files.sesync.org](files.sesync.org)).  

* Australia_NON-NATIVE_PLANT_FEEDERS_ants.xlsx
* Europe_Aug2019.xlsx
* Galapagos_Seebens_clean_only.xlsx
* Great Britain_Insect_pests_new_establishments_1970-2013.xlsx
* Hawaii_establishments_updateOct18.xlsx
* Japan_establishments_(confidential).xlsx
* Korea_established_species_from_Lee_etal_2016.xlsx
* New Zealand_Updated_exotic_insects_2019Oct11.xlsx
* North America_4.xlsx
* Ogasawara_establishments(confidential).xlsx
* Okinawa_establishments(Confidential).xlsx   

## Cleaning Scripts
Cleaning scripts are located on SESYNC GitLab in [example_code repository](https://gitlab.sesync.org/insectinvasions/example_code)

* custom_taxonomy_funcs.R
* make_taxonomy_table.R
* make_occurrence_table.R
* make_attribute_table.R

## Cleaned Data Tables
Tables are stored on the shared data directory at SESYNC (`/nfs/insectinvasions-data` or [files.sesync.org](files.sesync.org)).  
These data tables are the outputs of the cleaning scripts.  

* taxonomy_table.csv
* occurrence_table.csv
* attribute_table.csv

## Summary of data cleaning/table creation methods
1. Read in raw data from excel files for each region/country/area of interest using custom separate_taxonomy() function.  This 
function also cleans column names, and largely cleans up inconsistencies in species names.  The output is a dataframe.

2. Make two vectors, one of names to species and one of names to genus.  Individually send these vectors to the GBIF backbone 
taxonomy via the custom function get_accepted_taxonomy() to get the accepted taxonomic names.  

3. Take species names that only matched at genus level and try to get more info from other databases (**list other databases here)
for those taxa using custom get_more_info() function.  Also use this function to try and get info for species not found in GBIF. 

4. Read in CSV (`genus_only_resolution_FIXED.csv`) of manual fixes to species names, and merge this into the info from GBIF and 
other databases.  

5. Combine all accepted taxonomic names into one dataframe.  Add a new column with a unique id as a key (number for 1 to x).  Write 
dataframe to CSV file. 

6. For occurrence table, read in raw data from excel files for each area of interest using the custom separate_occurrence() 
function. This also cleans up country name and origin columns, and largely cleans up inconsistencies in species names.  

7. Join with the taxonomic table to add the unique taxon id column for key.  Write dataframe to CSV file. 

8. For attribute table, read in raw data from excel files for each area of interest using custom separate_attributes() function.  This
function cleans up column names, and largely cleans up inconsistencies in species names.   

9. Clean up origin names using regex.  Bring in taxonomic names by joining the taxonomic table.  

10. Bring in origin correspondence table (`origin_correspondence_table.xlsx`) to unify origin column entries.  Bring in non-plant-feeding info (`non-plant-feeding_taxa_updatedOct07.xlsx`) and join with this info in a new column.  Write dataframe to CSV file.  


## Advice on how to bring tables together
Note: All tables have Taxon ID column, which is unique ID for each cleaned taxa

1. read in taxonomy table and other table of interest
2. use a join to make the tables into one dataframe


```{r knit, echo=FALSE, message=FALSE, error=FALSE, eval=FALSE, include=FALSE}
ezknit(file = paste0(in_dir, "/Taxonomy_Cleaning_README.Rmd"), out_dir = out_dir, keep_html = FALSE)

render(paste0(out_dir, "/Taxonomy_Cleaning_README.md"), output_format = "pdf_document")
```

